How to compute the fusion product of MV cycles in type A

### 2021-04-21 draft.tex
still kind of a mess

\begin{proof}
    % Copy Roger's proof.
    We have
\[
\begin{split}
    S_{\mu_1, \mu_2} & = N\xT[t^{-1}]t^{\mu_1}(t-s)^{\mu_2} \\
     & \subset T_1\xt[t^{-1}] N\xT[t^{-1}] t^{\mu_1} (t-s)^{\mu_2} \\
     & = T_1\xt[t^{-1}] N_1\xt[t^{-1}] t^{\mu_1} (t-s)^{\mu_2} \qquad \text{\cite[Lemma 2.3]{kamnitzer2014yangians}}\\
     & = B_1\xt[t^{-1}] t^{\mu_1} (t-s)^{\mu_2} \\
     & = B_1\xt[t^{-1}] t^{\mu_1 + \mu_2} \\
     & \subset G_1\xt[t^{-1}] t^{\mu_1 + \mu_2} \\
     & = W_{\mu_1 + \mu_2}
\end{split}
\]
where $B_1\xt[t^{-1}] t^{\mu_1} (t-s)^{\mu_2} = B_1\xt[t^{-1}] t^{\mu_1 + \mu_2}$ since 
\[
\frac{t}{t-s} = 
1 + \frac{s}{t} + \frac{s^2}{t^2} + \cdots 
\in B_1\xt[t^{-1}].
\]
\end{proof}
\end{comment}
% 
% In this section we demonstrate how to put the main results of this paper into practice.
% 
Let $m,\lambda',\lambda'',\lambda,\mu',\mu'',\mu,N$ be as before.
% 
In this part we prove that 
% the ideal of 
$Z'\ast_\AA Z''$ can be computed from a ``minimal'' pair $(\tau',\tau'')$ of semistandard Young tableaux for the Lusztig data  $(n_\bullet(Z'),n_\bullet(Z''))$.
% 
Given $\lambda$ dominant effective and $\mu\le\lambda$ effective, let $YT(\lambda)_\mu$ denote the set of semistandard Young tableaux of shape $\lambda$ and weight $\mu$. 
%

We need the following result from \cite{dthesis}.
\begin{proposition}
    % Generalized orbital varieties are indexed by semistandard Young tableaux. 
    There are bijections 
    % Composing with the ordinary Mirkovi\'c--Vybornov isomorphism gives bijections 
    \[
        YT(\lambda)_\mu \to \irr \overline{\OO}^\lambda\cap\TT_\mu\cap\n \to \irr \overline\Gr^\lambda\cap S^\mu_-\,. 
    \]
    Moreover, there is a \acom{partial?} inverse 
    \[
    \CC[N]\to \bigcup_{\lambda~\text{dominant}\atop \mu\le\lambda} YT(\lambda)_\mu \,. 
    \]
    % $$
    % \begin{tikzcd}
    %     YT(\lambda)_\mu \ar[r] \ar[d] & \irr \overline{\OO}^\lambda\cap\TT_\mu\cap\n \ar[d] \\
    %     GT(\lambda)_\mu \ar[r] & \irr \overline\Gr^\lambda\cap S^\mu_-
    % \end{tikzcd}
    % $$
\end{proposition}
% 
\acom{Refer to Claxton--Tingley multisegments, Hong--Lee marignally large tableaux for comparison.}
\jcom{I don't think that you should mention this partial inverse here.  I think that you should explain here how the generalized orbital varieties are indexed by the tableaux, in other words, how $ X(\tau) $ is defined.}

Denote the image of $\tau$ in $\overline{\OO}^\lambda\cap\TT_\mu\cap\n$ by $X(\tau)$ and in $\overline{\Gr}^\lambda\cap S^\mu_-$ by $Z(\tau)$. Thus $Z(\tau) = \phi(X(\tau))$. \acom{Out of place?} 
\rcom{What is $\phi$?}

% 
% 
% TODO: Not yet, that would be the algo stuff that we are holding off on 
% In this section we explain how to determine the ideal of $Z'\ast_\AA Z''$ from the minimal pair $(\tau',\tau'')$ corresponding to $(n_\bullet(Z'),n_\bullet(Z''))$.
% 
% In particular, we prove that ``the generalized Spaltenstein map'' produces MV cycles. 
% 
% and recall the $\mu$-numeration of the standard basis of $\CC^N$. 
% TODO: can't we do without it^^
Given $A \in M_N(\CC)$ we'll denote by $A\big|_{\CC^i}$ the restriction of $A$ to the subspace spanned by the first $i$ standard basis vectors. Given $(\tau',\tau'')\in YT(\lambda')_{\mu'}\times YT(\lambda'')_{\mu''}$ denote by $\lambda'(i),\lambda''(i)$ the shape of $\tau'(i),\tau''(i)$ resp.\ where $\tau^?(i)$ denotes the tableau got from $\tau^?$ by discarding all but its first $i$ boxes. If the tableau contains repeated entries then first means leftmost.

\begin{proposition}
    Given $Z',Z''$, let $(\tau',\tau'')\in YT(\lambda')_{\mu'}\times YT(\lambda'')_{\mu''}$ be such that\dots If $s\ne 0$, then 
    \[
    \overline{\{A\in\UU_{0,s}^{\mu',\mu''} : A \big|_{\CC^{i}} \in \OO_{0,s}^{\lambda'(i),\lambda''(i)} \text{ all } i=1,\dots,N\}} 
    \cong Z'\times Z'' \,. %\mu_1 + \cdots + \mu_i
    % = Z(\tau')\times Z(\tau'')\,. 
    \]
\end{proposition}
% 
\jcom{Do you want the ``blocky'' or ``boxy'' description here?  I mean, do you want $ i = 1, \dots, N$ or $ i =1, \dots, m$? I'm not so sure that ``boxy'' will work, since with those $s$ in the matrix, it doesn't preserves all the $ \CC^i$.}

\begin{proof}
    Let $(A,s)$ be an element of the set on the left hand side, let $a\in\{0,s\}$ and denote by $E_a$ the generalized $a$-eigenspace of $A$. 
    Let $i\in \{1,\dots,N\}$ and write $m(i)$ for the maximum weight of a box in $\tau'(i)$ or $\tau''(i)$.

    The the Jordan type of $A-a\big|_{E_a\cap\CC^{m(i)}}$ is given by $\lambda'(i)$ if $a = 0$ and $\lambda''(i)$ if $a = s$. 
    Moreover, the lattice $(L,s)$ defined by $g(A)$ is such that $t-a$ has the same Jordan type on 
    % $\cO_a^{m(i)} / L(a)\cap\cO_a^{m(i)} = \CC[t]^{m(i)}/L\cap\CC[t]^{m(i)}\otimes_{\CC[t]}\cO_a $ 
    \begin{equation}
        \label{eq:bcflat}
        \cO_a^{m(i)} / L(a)\cap\cO_a^{m(i)} = \CC[t]^{m(i)}/L\cap\CC[t]^{m(i)}\otimes_{\CC[t]}\cO_a 
    \end{equation}
    as does $(A-a\big|_{E_a\cap\CC^{m(i)}},s)$. 
    \acom{Not sure when to use $(?,s)$ notation and when not in this proof. Here I just didn't want to end the sentence with the bulky subscript.}
    % 
    \Cref{eq:bcflat} follows from e.g.\ \cite[Proposition~10.14]{atiyah2018introduction}:
    \begin{quotation}
        If $A$ is a Noetherian ring, $\fa$ an ideal, $\hat A$ the $\fa$-adic completion of $A$, then $\hat A$ is a flat $A$-algebra.
    \end{quotation}
    applied to $A = \CC[t]$ and $\fa = (t)$.

@@

    The second assertion is a consequence of \Cref{lem:} \acom{Cite Joel's lemma}. 
    By \Cref{adthesis} \acom{Cite my thesis} it follows that $(L(0),L(s)) \in Z'\times Z''$. 
    Since all maps are isomorphisms the converse is automatic. 
\end{proof}

@@

\begin{lemma}
    \label{lem:mintab}
    Let $n_\bullet$ be a Lusztig datum of coweight $\nu$. There exist $\lambda,\mu$ smallest such that $n_\bullet$ is the Lusztig datum of a tableau $\tau\in YT(\lambda)_\mu$. They are determined by the following minimization problem. Regard $n_\bullet$ as a strictly lower triangular matrix with $(r,c)$th entry $n_{r,c}$ is equal to the number of boxes of weight $r$ in row $c$ of $\tau$. Note that since $r > c$ this data does not account for the number of boxes of weight $r$ in row $r$. Let $\mu^0$ be the diagonal matrix whose $(r,r)$th entry $\mu^0_r$ is equal to the number of boxes of weight $r$ in row $r$ of $\tau$. Note that the sum of the rows of $n_\bullet + \mu^0$ is the shape $\lambda$ of $\tau$, and the sum of the columns of $n_\bullet + \mu^0$ is the weight $\mu$ of $\tau$. The solution to the minimization problem 
    \[
        \begin{gathered}
        % 
        \text{minimize }\mu^0\text{ such that} \\
        % \text{the sum of the rows of }n_\bullet + \mu^0\text{ is non-increasing},
        % \text{the sum of the columns of }n_\bullet + \mu^0\text{ is non-increasing}
        \text{both }\lambda,\mu\text{ are non-increasing}
        \end{gathered} 
    \]
    is got recursively from % setting $\mu^0_m = 0$, 
    \[
        \begin{gathered}
            \mu^0_{m} = 0 \\ 
            \mu^0_{m-1} = \max\{0 , \sum n_{m,c} - \sum n_{m-1,c}\} \\ 
            \mu^0_i = \max\{0, \mu^0_{i+1} + \sum n_{i+1,c} - \sum n_{i,c} , \mu_{i+1}^0 + \sum n_{r,i+1} - \sum n_{r,i}\}
        \end{gathered}
    \]
    for $i = 1,\dots,m-2$. 
    In words, given a Luszig datum $n_\bullet$, if we wish to produce a minimal SSYT for it, 
    \begin{itemize}
        \item we can always take the number of $m$'s in row $m$ to be zero; 
        \item we can take the number of $m-1$'s in row $m-1$ to be zero, unless $n_\bullet$ tells us that there are more $m$'s the first $m-1$ rows than there are $m-1$'s in the first $m-2$ rows, and in that case we take $\mu^0_{m-1}$ so that it exactly offsets the difference; 
        \item we can take the number of $m-2$'s in row $m-2$ to be zero, unless $n_\bullet$ tells us there are more $m-1$'s in the first $m-1$ rows (building off the last step) than there are $m-2$'s in the first $m-3$ rows, or, that there are more boxes in row $m-1$ than there are in row $m-2$, and then we take $\mu^0_{m-2}$ to offset the minimum of the two differences; 
        \item induct up to $\mu^0_1$. 
    \end{itemize}
\end{lemma}

>>>>

\begin{proof}
    % Copy Roger's proof.
    We have
\[
\begin{split}
    S_{\mu_1, \mu_2} & = N\xT[t^{-1}]t^{\mu_1}(t-s)^{\mu_2} \\
     & \subset T_1\xt[t^{-1}] N\xT[t^{-1}] t^{\mu_1} (t-s)^{\mu_2} \\
     & = T_1\xt[t^{-1}] N_1\xt[t^{-1}] t^{\mu_1} (t-s)^{\mu_2} \qquad \text{\cite[Lemma 2.3]{kamnitzer2014yangians}}\\
     & = B_1\xt[t^{-1}] t^{\mu_1} (t-s)^{\mu_2} \\
     & = B_1\xt[t^{-1}] t^{\mu_1 + \mu_2} \\
     & \subset G_1\xt[t^{-1}] t^{\mu_1 + \mu_2} \\
     & = W_{\mu_1 + \mu_2}
\end{split}
\]
where $B_1\xt[t^{-1}] t^{\mu_1} (t-s)^{\mu_2} = B_1\xt[t^{-1}] t^{\mu_1 + \mu_2}$ since 
\[
\frac{t}{t-s} = 
1 + \frac{s}{t} + \frac{s^2}{t^2} + \cdots 
\in B_1\xt[t^{-1}].
\]
\end{proof}
\end{comment}
% 
We begin this section with some new notation. Let $YT(\lambda)_\mu$ denote the set of (possibly semistandard) Young tableaux of (arbitrary) shape $\lambda$ and (arbitrary) weight $\mu$. Here, the weight of a tableau is a sequence $(\mu_1,\mu_2,\dots)$ recording the number of boxes of weight $i = 1,2,\dots$ that the tableau contains, and the weight of a box is just the number that the box contains.
% For the sake of making a definition, we momentarily put some letters of our alphabet in unexpected roles, and ask you to suspend your disbelief. Given an arbitrary tableau $t$ having weight $w$ and shape $s$, we denote by $s(i)$ (resp.\ $w(i)$) the shape (resp.\ the weight) of $t(i)$, the tableau got from $t$ by discarding all boxes of weight exceeding $i$.

Given $\tau\in YT(\lambda)_\mu$ we denote by $\lambda(i)$ (resp.\ $\mu(i)$) the shape (resp.\ the weight) of $\tau(i)$, the tableau got from $\tau$ by discarding all boxes of weight exceeding $i$.

Given $A \in M_N(\CC)$ we denote by 
% $A\big|_{\CC^i}$ 
$A\big|_{\CC^p}$ 
the restriction of $A$ to the subspace spanned by the first $p$ standard basis vectors. If $A\big|_{\CC^p}$ preserves $\CC^p$ then we identify it with the $p\times p$ upper-left submatrix of $A$.

The variety $X(\tau)$ defined by
% from $\tau\in YT(\lambda)_\mu$ by 
\[
X(\tau) = \overline{
    \{
        A \in  \overline{\OO}^\lambda\cap\TT_\mu\cap\n : A\big|_{\CC^{|\mu(i)|}} \in \OO^{\lambda(i)} \text{ for each } i = 1,\dots,m
    \}
}    
\]
is called a generalized orbital variety.

% 
We need the following result from \cite{dthesis}.
\begin{proposition}
    \label{prop:dth}
    (Part of which is a corollary of \cite[Theorem~4.4.1(c)]{mirkovic2019comparison})
    There is an isomorphism 
    $\phi: \overline{\OO}^\lambda\cap\TT_\mu\cap\n \to \overline\Gr^\lambda\cap S^\mu_-$ 
    and it is compatible with the bijection $YT(\lambda)_\mu \to \irr \overline{\OO}^\lambda\cap\TT_\mu\cap\n$ 
    in the sense that the irreducible component $\phi(X(\tau))$ associated to the semistandard Young tableau $\tau$ is an open dense subset of the MV cycle whose Lusztig datum is equal to that of $\tau$. 
\end{proposition}

% Given $(\tau',\tau'')\in YT(\lambda')_{\mu'}\times YT(\lambda'')_{\mu''}$ denote by $\lambda'(i),\lambda''(i)$ (resp.\ $\mu'(i),\mu''(i)$) the shape (resp.\ the weight) of $\tau'(i),\tau''(i)$ where $\tau^?(i)$ denotes the tableau got from $\tau^?$ by discarding all 
% % but its first $i$ boxes
% boxes of weight exceeding $i$. 
% % If the tableau contains repeated entries then first means leftmost.
% 
\jcom{I don't think that you should mention this partial inverse here.  I think that you should explain here how the generalized orbital varieties are indexed by the tableaux, in other words, how $ X(\tau) $ is defined.}
\acom{Ok, I moved it.}

% Denote the image of $\tau$ in $\overline{\OO}^\lambda\cap\TT_\mu\cap\n$ by $X(\tau)$ and in $\overline{\Gr}^\lambda\cap S^\mu_-$ by $Z(\tau)$. Thus $Z(\tau) = \phi(X(\tau))$. \acom{Out of place?} 
% \rcom{What is $\phi$?}
% 
% 
% TODO: Not yet, that would be the algo stuff that we are holding off on 
% In this section we explain how to determine the ideal of $Z'\ast_\AA Z''$ from the minimal pair $(\tau',\tau'')$ corresponding to $(n_\bullet(Z'),n_\bullet(Z''))$.
% 
% In particular, we prove that ``the generalized Spaltenstein map'' produces MV cycles. 
% 
% and recall the $\mu$-numeration of the standard basis of $\CC^N$. 
% TODO: can't we do without it^^

% 
% In this section we demonstrate how to put the main results of this paper into practice.
% 
Now let $m,\lambda',\lambda'',\lambda,\mu',\mu'',\mu,N$ be as before.
% 
In this part we prove that 
% the ideal of 
$Z'\ast_\AA Z''$ can be computed from a ``minimal'' pair $(\tau',\tau'')$ of semistandard Young tableaux for the Lusztig data  $(n_\bullet(Z'),n_\bullet(Z''))$. 
\acom{Need to introduce LD.}
\begin{proposition}
    Given $b_{Z'},b_{Z''}\in\CC[N]$, let $(\tau',\tau'')\in YT(\lambda')_{\mu'}\times YT(\lambda'')_{\mu''}$ be such that $t^{\lambda'}Z' = X(\tau')$ and $t^{\lambda''}Z'' = X(\tau'')$. Let $\mu = \mu' + \mu''$. If $s\ne 0$, then 
    \[
    \overline{\{A\in\UU_{0,s}^{\mu',\mu''} : A \big|_{\CC^{|\mu(i)|}} \in \OO_{0,s}^{\lambda'(i),\lambda''(i)} \text{ all } i=1,\dots,N\}} 
    \cong Z'\times Z'' \,. %\mu_1 + \cdots + \mu_i
    % = Z(\tau')\times Z(\tau'')\,. 
    \]
\end{proposition}
% 
\jcom{Do you want the ``blocky'' or ``boxy'' description here?  I mean, do you want $ i = 1, \dots, N$ or $ i =1, \dots, m$? I'm not so sure that ``boxy'' will work, since with those $s$ in the matrix, it doesn't preserves all the $ \CC^i$.}

\acom{Fixed.}

\begin{proof}
    Let $(A,s)$ be an element of the set on the left hand side, let $a\in\{0,s\}$ and denote by $E_a$ the generalized $a$-eigenspace of $A$. 
    % Let $i\in \{1,\dots,N\}$ and write $m(i)$ for the maximum weight of a box in $\tau'(i)$ or $\tau''(i)$.

    The the Jordan type of $A-a\big|_{E_a\cap\CC^{|\mu(i)|}}$ is given by $\lambda'(i)$ if $a = 0$ and $\lambda''(i)$ if $a = s$. 
    Moreover, the lattice $(L,s)$ defined by $g(A)$ is such that $t-a$ has the same Jordan type on 
    % $\cO_a^{m(i)} / L(a)\cap\cO_a^{m(i)} = \CC[t]^{m(i)}/L\cap\CC[t]^{m(i)}\otimes_{\CC[t]}\cO_a $ 
    \begin{equation}
        \label{eq:bcflat}
        \cO_a^{|\mu(i)|} / L(a)\cap\cO_a^{|\mu(i)|} = \CC[t]^{|\mu(i)|}/L\cap\CC[t]^{|\mu(i)|}\otimes_{\CC[t]}\cO_a 
    \end{equation}
    as does $(A-a\big|_{E_a\cap\CC^{|\mu(i)|}},s)$. 
    \acom{Not sure when to use $(?,s)$ notation and when not in this proof. Here I just didn't want to end the sentence with the bulky subscript.}
    % 
    % \Cref{eq:bcflat} follows from e.g.\ \cite[Proposition~10.14]{atiyah2018introduction}:
    % \begin{quotation}
    %     If $A$ is a Noetherian ring, $\fa$ an ideal, $\hat A$ the $\fa$-adic completion of $A$, then $\hat A$ is a flat $A$-algebra.
    % \end{quotation}
    % applied to $A = \CC[t]$ and $\fa = (t)$. 
    \Cref{eq:bcflat} follows from the fact that a module over a PID is flat if and only if it is torsion free, and $\cO_a$ is clearly torsion free.

@@

    The second assertion is a consequence of \Cref{le:linalg}.
    By \Cref{prop:dth}, it follows that $(L(0),L(s)) \in Z'\times Z''$. 
    Since all maps are isomorphisms the converse is automatic. 
\end{proof}

@@

\acom{How's here?}
It may be helpful to know the map
\[
    \CC[N]\to \bigcup_{\lambda~\text{dominant}\atop \mu\le\lambda} YT(\lambda)_\mu 
\]
that we use to find our tableaux.
    % $$
    % \begin{tikzcd}
    %     YT(\lambda)_\mu \ar[r] \ar[d] & \irr \overline{\OO}^\lambda\cap\TT_\mu\cap\n \ar[d] \\
    %     GT(\lambda)_\mu \ar[r] & \irr \overline\Gr^\lambda\cap S^\mu_-
    % \end{tikzcd}
    % $$
\acom{Refer to Claxton--Tingley multisegments, Hong--Lee marignally large tableaux for comparison.}
\begin{lemma}
    \label{lem:mintab}
    Let $n_\bullet$ be a Lusztig datum of coweight $\nu$. There exist $\lambda,\mu$ smallest such that $n_\bullet$ is the Lusztig datum of a tableau $\tau\in YT(\lambda)_\mu$. They are determined by the following minimization problem. Regard $n_\bullet$ as a strictly lower triangular matrix with $(r,c)$th entry $n_{r,c}$ is equal to the number of boxes of weight $r$ in row $c$ of $\tau$. Note that since $r > c$ this data does not account for the number of boxes of weight $r$ in row $r$. Let $\mu^0$ be the diagonal matrix whose $(r,r)$th entry $\mu^0_r$ is equal to the number of boxes of weight $r$ in row $r$ of $\tau$. Note that the sum of the rows of $n_\bullet + \mu^0$ is the shape $\lambda$ of $\tau$, and the sum of the columns of $n_\bullet + \mu^0$ is the weight $\mu$ of $\tau$. The solution to the minimization problem 
    \[
        \begin{gathered}
        % 
        \text{minimize }\mu^0\text{ such that} \\
        % \text{the sum of the rows of }n_\bullet + \mu^0\text{ is non-increasing},
        % \text{the sum of the columns of }n_\bullet + \mu^0\text{ is non-increasing}
        \text{both }\lambda,\mu\text{ are non-increasing}
        \end{gathered} 
    \]
    is got recursively from % setting $\mu^0_m = 0$, 
    \[
        \begin{gathered}
            \mu^0_{m} = 0 \\ 
            \mu^0_{m-1} = \max\{0 , \sum n_{m,c} - \sum n_{m-1,c}\} \\ 
            \mu^0_i = \max\{0, \mu^0_{i+1} + \sum n_{i+1,c} - \sum n_{i,c} , \mu_{i+1}^0 + \sum n_{r,i+1} - \sum n_{r,i}\}
        \end{gathered}
    \]
    for $i = 1,\dots,m-2$. 
    In words, given a Luszig datum $n_\bullet$, if we wish to produce a minimal SSYT for it, 
    \begin{itemize}
        \item we can always take the number of $m$'s in row $m$ to be zero; 
        \item we can take the number of $m-1$'s in row $m-1$ to be zero, unless $n_\bullet$ tells us that there are more $m$'s the first $m-1$ rows than there are $m-1$'s in the first $m-2$ rows, and in that case we take $\mu^0_{m-1}$ so that it exactly offsets the difference; 
        \item we can take the number of $m-2$'s in row $m-2$ to be zero, unless $n_\bullet$ tells us there are more $m-1$'s in the first $m-1$ rows (building off the last step) than there are $m-2$'s in the first $m-3$ rows, or, that there are more boxes in row $m-1$ than there are in row $m-2$, and then we take $\mu^0_{m-2}$ to offset the minimum of the two differences; 
        \item induct up to $\mu^0_1$. 
    \end{itemize}
\end{lemma}
